Equalizer Programming Guide

### 2007-08-21 Developer/ProgrammingGuide/paper.tex
Update Programming Guide
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\usepackage{tabularx,graphicx,a4}

@@

\title{Equalizer Programming Guide}
\author{Stefan Eilemann\thanks{eile@eyescale.ch}\\[\medskipamount]
  % Eyescale Software Sarl
}
\date{
  \textbf{INCOMPLETE}\\[\medskipamount]
  Version 0.1, \today
}

@@

\maketitle
\vfill

@@

\vfill {\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Version & \bf Date     & \bf Changes \\
    \hline
    0.1         & Aug 19, 2007 & outlined the basic concepts\\
    \hline \multicolumn{3}{c}{\small
      \htmladdnormallink{http://www.equalizergraphics.com/documents/Developer/ProgrammingGuide.pdf}
      {http://www.equalizergraphics.com/documents/ProgrammingGuide.pdf}}\\
  \end{tabularx}}

@@

The application subclasses Equalizer objects and overrides virtual
functions to implement certain functionality, e.g., the OpenGL rendering
code in \textsf{eq::Channel::frameDraw}. These task methods are in
concept similar to C function callbacks.

\section{The Resource Tree}

@@

\subsection{Configuration}

@@

\subsection{Node}

@@

\subsection{Pipe}

The \textsf{eq::Pipe} is the abstraction of a graphics card (GPU). In
the current implementation it is also one operating system thread,
unless the pipe's thread hint is set to \textsf{false}. All pipe and
child window and channel task methods are executed from the pipe thread
for threaded pipes or from the main application thread for non-threaded
pipes\footnote{see also
  \htmladdnormallink{http://www.equalizergraphics.com/documents/design/nonthreaded.html}{http://www.equalizergraphics.com/documents/design/nonthreaded.html}
  for non-threaded rendering}.

@@

\subsection{Window}

@@

\subsection{Channel}

@@

\bibliographystyle{abbrvurl}
\bibliography{paper}
\end{document}

>>>>

\usepackage{tabularx,graphicx,a4,listings}

@@

\title{Equalizer Programming Guide}
\author{Stefan Eilemann\thanks{eile@eyescale.ch}\\[\medskipamount]
  % Eyescale Software Sarl
}
\date{
  \textbf{INCOMPLETE}\\[\medskipamount]
  Version 0.2, \today
}

@@

\maketitle
\vfill
\lstset{language=C++}

@@

\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Version & \bf Date     & \bf Changes \\
    \hline
    0.2         & Aug 20, 2007 & main function described\\
    0.1         & Aug 19, 2007 & outlined the basic concepts\\
    \hline \multicolumn{3}{c}{\small
      \htmladdnormallink{http://www.equalizergraphics.com/documents/Developer/ProgrammingGuide.pdf}
      {http://www.equalizergraphics.com/documents/ProgrammingGuide.pdf}}\\
  \end{tabularx}}

@@

The application subclasses Equalizer objects and overrides virtual
functions to implement certain functionality, e.g., the application's
OpenGL rendering in \textsf{eq::Chan\-nel::frameDraw}. These task
methods are in concept similar to C function callbacks. The
\textsf{eqPly} section will discuss the most important task methods. A
full list of all task methods can be found on the
website\footnote{\htmladdnormallink{http://www.equalizergraphics.com/documents/design/taskMethods.html}{http://www.equalizergraphics.com/documents/design/taskMethods.html}}.

\subsection{The Resource Tree}

@@

\subsubsection{Configuration}

@@

\subsubsection{Node}

@@

\subsubsection{Pipe}

The \textsf{eq::Pipe} is the abstraction of a graphics card (GPU). In
the current implementation it is also one operating system thread,
unless the pipe's thread hint is set to \textsf{false}. All pipe and
child window and channel task methods are executed from the pipe thread
for threaded pipes or from the main application thread for non-threaded
pipes\footnote{see also
  \htmladdnormallink{http://www.equalizergraphics.com/documents/design/nonthreaded.html}{http://www.equalizergraphics.com/documents/design/nonthreaded.html}}.

@@

\subsubsection{Window}

@@

\subsubsection{Channel}

@@

\subsection{Resource Usage}

How the rendering resources are to be used is configured using a
compound tree. Each compound has a channel, which it uses to execute the
rendering tasks. The rendering tasks are computed by the server and send
to the render clients. At no point the application or render clients
have or need knowledge of compounds. The configuration of compounds is
not in the scope of this document\footnote{see
  \htmladdnormallink{http://www.equalizergraphics.com/documents/design/compounds.html}{http://www.equalizergraphics.com/documents/design/compounds.html}}.

\section{The \textsf{eqPly} polygonal renderer}

The \textsf{eqPly} example is shipped with the Equalizer distribution
and serves as a simple reference implementation of an Equalizer-based
application. Its focus is not on rendering features or visual quality.
It serves as a test bed for most of the Equalizer features.

In this section the source code of \textsf{eqPly} is discussed in
detail, and relevant design decision and remarks are raised.

All classes in the example are in the \textsf{eqPly} namespace to avoid
type name ambiguities, in particular for the \textsf{Window} class.

\subsection{The main Function}

The main function starts off with parsing the command line into the
\textsf{LocalInitData} data structure, which in part will be distributed
to all render client nodes. For actual command line parsing is done by
the \textsf{LocalInitData} class and will be discussed there:

{\small\begin{lstlisting}
int main( int argc, char** argv )
{
    // 1. parse arguments
    eqPly::LocalInitData initData;
    initData.parseArguments( argc, argv );
\end{lstlisting}}

The second step is to initialize the Equalizer library. The
initialization function of Equalizer also parses the command line, which
is used to set certain default values based on Equalizer-specific
options\footnote{Equalizer-specific options always start with --eq-},
e.g., the default server location. Furthermore, a node factory is
provided:

{\small\begin{lstlisting}
    // 2. Equalizer initialization
    NodeFactory nodeFactory;
    if( !eq::init( argc, argv, &nodeFactory ))
    {
        EQERROR << "Equalizer init failed" << endl;
        return EXIT_FAILURE;
    }
\end{lstlisting}}

The node factory is used by Equalizer to create the object instances for
the rendering entities. Each of the classes inherits from the same type
provided by Equalizer in the \textsf{eq} namespace. The provided
\textsf{eq::NodeFactory} base class instantiates a 'plain' Equalizer
object, thus making it possible to selectively subclass individual
entity types. For each rendering resource used in the configuration, one
C++ object will be created:

{\small\begin{lstlisting}
class NodeFactory : public eq::NodeFactory
{
public:
    virtual eq::Config*  createConfig()  { return new eqPly::Config; }
    virtual eq::Node*    createNode()    { return new eqPly::Node; }
    virtual eq::Pipe*    createPipe()    { return new eqPly::Pipe; }
    virtual eq::Window*  createWindow()  { return new eqPly::Window; }
    virtual eq::Channel* createChannel() { return new eqPly::Channel; }
};
\end{lstlisting}}

The third step is to create an instance of the application and to
initialize it locally. The application is an \textsf{eq::Client}, which
is an \textsf{eqNet::Node}. The underlying network distribution in
Equalizer is a peer-to-peer network structure of
\textsf{eqNet::Node}s. The application programmer rarely is aware of the
classes in the \textsf{eqNet} namespace, but both the
\textsf{eq::Client} and the server are \textsf{eqNet::Node}s. The local
initialization of nodes creates a local listening socket, so that the
node, and therefore the \textsf{eq::Client} can communicate over the
network with other nodes, such as the server and the rendering clients.

{\small\begin{lstlisting}
    // 3. initialization of local client node
    RefPtr< eqPly::Application > client = new eqPly::Application( initData );
    if( !client->initLocal( argc, argv ))
    {
        EQERROR << "Can't init client" << endl;
        eq::exit();
        return EXIT_FAILURE;
    }
\end{lstlisting}}

Finally everything is set up to run the \textsf{eqPly} application:

{\small\begin{lstlisting}
    // 4. run client
    const int ret = client->run();
\end{lstlisting}}

After it has finished, the application and Eqply is deinitialized and
the \textsf{main} function returns:

{\small\begin{lstlisting}
    // 5. cleanup and exit
    client->exitLocal();
    client = 0;

    eq::exit();
    return ret;
}
\end{lstlisting}}

\subsection{Application}

In the \textbf{eqPly} case, the application is also the render client. It has
three run-time behaviours:

\begin{enumerate}
  \item \textbf{Application}: The executable started by the user, which
    is the controlling entity in the rendering session.
  \item \textbf{Auto-launched Render Client}: The typical render client,
    started by the server. In this mode, \textsf{initLocal} never
    returns and the process is terminated by the server.
  \item \textbf{Resident Render Client}: Manually pre-started render
    client, listening on a specified port for server commands.
\end{enumerate}

\end{document}
